---
title: "Analysis of Flights"
output:
  pdf_document: default
  html_document:
    df_print: paged
fontsize: 12pt
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
I will be analyzing the flights data set and answering various prompts


### Libraries

```{r}
library(dplyr)
library(readr)
library(memoise)
```


### Data set

```{r}
flights = as_tibble(data.table::fread("https://github.com/Rdatatable/data.table/blob/master/vignettes/flights14.csv?raw=true"))
flights
```

### Analysis
This section will mainly involve demonstrating an understanding of different aspects of dplyr rather than focusing on a prompt.

Finding the average dep_delay by carrier
```{r}
flights %>%
  group_by(carrier) %>%
  summarise(avg_dep_delay = mean(dep_delay))
```

Finding total delay of each flight
```{r}
flights %>%
  mutate(tot_delay = arr_delay + dep_delay) %>%
  select(carrier, origin, tot_delay)
```
Top 5 origin-dest pairs with most flights?

```{r}
flights %>%
  group_by(origin, dest) %>%
  summarise(N = n()) %>%
  arrange(desc(N)) %>%
  head(5)
```
Average departure and arrival delay by month

```{r}
flights %>%
  group_by(month) %>%
  summarise(avg_dep_delay = mean(dep_delay),
            avg_arr_delay = mean(arr_delay))
```

Which days of the month have the most on-time arrivals?

```{r}
flights %>%
  group_by(day) %>%
  filter(arr_delay <= 0) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
```

Delay percentiles by carrier

```{r}
flights %>%
  group_by(carrier) %>%
  summarise(arr_delay_quan = quantile(arr_delay)) # fix later
```



