---
title: "Measuring Student Performance"
output:
  pdf_document: default
  html_document:
    df_print: paged
fontsize: 12pt
---

### Libraries

```{r, echo = FALSE}
library(tidyverse)
library(shiny)
library(ggplot2)
library(leaps)
```


### Dataset

```{r, echo = FALSE}
performance = read_csv("https://raw.githubusercontent.com/austinkirwin/public-projects/refs/heads/main/R_projects/StudentsPerformanceProject/StudentsPerformance.csv")
head(performance)
```


### Analysis

## Intro
How much does a student's upbringing affect their test scores? Does their race matter? Gender? Is it possible that these factors might be intertwined in affecting a student's result? This analysis hopes to answer some of those questions. The Student Performance data set was found on [Kaggle](https://www.kaggle.com/datasets/spscientist/students-performance-in-exams) and does not include any deeper explanations on the included variables (what race "groups" are, what the test preparation consisted of, etc). 

## Data Exploration

One-Hot Encoding categorical variables:
```{r, echo=FALSE}
encodedgender = model.matrix(~gender-1, data = performance)

# Renaming columns so they are easier to work with
lookup = c(race = "race/ethnicity", parentEdu = "parental level of education", testPrep = "test preparation course", mathScore = "math score", readingScore = "reading score", writingScore = "writing score")
performance = rename(performance, all_of(lookup))

encodedrace = model.matrix(~race-1, data = performance)
encodedparent = model.matrix(~parentEdu-1, data = performance)
encodedlunch = model.matrix(~lunch-1, data = performance)
encodedtest = model.matrix(~testPrep-1, data = performance)

performance_df = as.data.frame(cbind(encodedgender, encodedrace, encodedparent, encodedlunch, encodedtest, performance$mathScore, performance$readingScore, performance$writingScore))

# Renaming columns in new data frame
newNames = c(raceA = "racegroup A", raceB = "racegroup B", raceC = "racegroup C", raceD = "racegroup D", raceE = "racegroup E", parentEduAssoc = "parentEduassociate's degree", parentEduBach = "parentEdubachelor's degree", parentEduHS = "parentEduhigh school", parentEduMaster = "parentEdumaster's degree", parentEduSomeCollege = "parentEdusome college", parentEduSomeHS = "parentEdusome high school", lunchFree = "lunchfree/reduced", mathScore = "V18", readingScore = "V19", writingScore = "V20")
performance_df = rename(performance_df, all_of(newNames))
head(performance_df)
```
Checking for NA values:
```{r, echo=FALSE}
dim(performance_df) - dim(na.omit(performance_df))
```
The size of the dataframe with NA values is [0,0] so there are no NA values.

```{r}
test = c()
a = 1
for (i in 1:nrow(abs(cor(performance_df)))) {
  for (j in 1:ncol(abs(cor(performance_df)))) {
    if (abs(cor(performance_df))[i, j] <= .6) {
      test[a] == abs(cor(performance_df))[i, j]
      a = a+1
    }
  }
}  
test
```
## Full Model

Full model, reference group not shown in output
```{r, echo = FALSE}
g = lm(mathScore ~ ., data = performance)
g
```
Best subsets of model:
```{r}
a= regsubsets(mathScore~., data = performance)
rs = summary(a)
rs$which
```
```{r}
step(g, direction = "both")
```






